---
title: "Broad vs. Sanger Cell Line IDs"
author: "Ariel Balter"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: yes
    theme: lumen
    fig_caption: yes
    code_folding: show
---


```{r include=FALSE}
library(tidyverse)
library(DT)

display_table = function(
    df,
    caption=NULL,
    search=list(regex = TRUE, caseInsensitive = TRUE)
)
{
  num_rows = min(10, nrow(df))
  df %>%
    datatable(
      extensions = c('ColReorder', 'Scroller'),
      options = list(
        colReorder=T,
        scroller=T,
        scrollY=20*num_rows,
        ScrollX=T,
        ScrollY=T,
        autoWidth=T,
        searchHighlight = TRUE,
        search = search,
        pageLength = num_rows,
        fixedheader = TRUE
      ),
      caption = htmltools::tags$caption(
        style = '
        caption-side: top;
        text-align: center;
        color:black;
        font-size:150% ;',
        caption),
      # caption = caption,
      rownames = F
    ) %>%
    formatStyle(columns = colnames(.), fontSize = '50%')
}
```


# DepMap Project
Broad and Sanger are both part of the DepMap project. 

## Sanger
Sanger has a web page for its [DepMap Models](https://depmap.sanger.ac.uk/programmes/#models
)

Under this section is a link to the [Cell Model Passports](https://cellmodelpassports.sanger.ac.uk/
) section which

>provides a single location where information on Sanger DepMap cell models is available in a user-friendly environment.

Cell Model Passports has a [download page](https://cellmodelpassports.sanger.ac.uk/downloads) which provides 

>Stable [link](https://cog.sanger.ac.uk/cmp/download/model_list_latest.csv.gz) that always points to the latest version.

## Broad
Broad hosts data for the DepMap project at a dedicated portal:

https://depmap.org/portal/download/

Broad also has a seemingly-related project called the [Cancer Cell Line Encyclopedia (CCLE)](https://sites.broadinstitute.org/ccle. The [CCLE Datasets page](https://sites.broadinstitute.org/ccle/datasets) has a link for an annotated list of cell lines, however, that link is dead. The link for Processed Data leads to the DepMap download portal.

That portal lists a file called [_sample_info.csv_](https://ndownloader.figshare.com/files/35020903
) which could very well be the annotated cell line information.



# Download Sanger Model List
```{r label="load wide tables"}
model_list = 
  read_csv("https://cog.sanger.ac.uk/cmp/download/model_list_latest.csv.gz") %>% 
  select(
    sanger_model_id = model_id,
    depmap_id = BROAD_ID,
    sanger_sample_id = sample_id,
    sanger_patient_id = patient_id,
    model_type,
    cell_line_name = model_name,
    ccle_id = CCLE_ID,
    tissue,
    cancer_type,
    cancer_subtype = cancer_type_detail,
    sample_site
  ) %>% 
  mutate(from_sanger = T) %>% 
  rowwise() %>% 
  mutate(depmap_id = str_split(depmap_id, ";")) %>% 
  unnest_longer(depmap_id)
```

# Download BROAD DepMap "Sample Info"

```{r}
sample_info = 
  read_csv("https://ndownloader.figshare.com/files/35020903") %>% 
  select(
    depmap_id = DepMap_ID,
    sanger_model_id = Sanger_Model_ID,
    ccle_id = CCLE_Name,
    cell_line_name,
    stripped_cell_line_name,
    tissue = sample_collection_site,
    cancer_type = primary_disease,
    cancer_subtype = Subtype,
    lineage,
    lineage_subtype
  ) %>% 
  mutate(from_broad = T)

```

# Joined
```{r}
joined  = 
  full_join(
    sample_info,
    model_list,
    by = c("sanger_model_id", "depmap_id"),
    suffix = c("_broad", "_sanger")
  ) %>% 
  distinct() %>% 
  mutate(
    has_depmap_id = !is.na(depmap_id),
    has_sanger_id = !is.na(sanger_model_id),
    has_ccle_id_broad = !is.na(ccle_id_broad),
    has_ccle_id_sanger = !is.na(ccle_id_sanger),
    has_cell_line_name_broad = !is.na(cell_line_name_broad),
    has_cell_line_name_sanger = !is.na(cell_line_name_sanger)
  ) %>% 
  group_by(depmap_id) %>% 
    filter(n()==1 | n()>1 & !is.na(sanger_model_id)) %>% 
  ungroup() %>% 
  group_by(sanger_model_id) %>% 
    filter(n()==1 | n()>1 & !is.na(depmap_id)) %>% 
  ungroup()

sorted_colnames = 
  colnames(joined) %>% 
  sort() %>% 
  setdiff(., c("sanger_model_id", "depmap_id")) %>% 
  c(c("sanger_model_id", "depmap_id"), .)

joined = joined %>% select(!!sorted_colnames)

joined %>% display_table()

```


# Some Counts
```{r}
joined %>% 
  mutate(
    has_depmap_id = !is.na(depmap_id),
    has_sanger_id = !is.na(sanger_model_id)
  ) %>% 
  count(has_depmap_id, has_sanger_id) %>%
  display_table()

joined %>% 
  mutate(
    has_depmap_id = !is.na(depmap_id),
    has_sanger_id = !is.na(sanger_model_id)
  ) %>% 
  group_by(model_type) %>% 
  count(has_depmap_id, has_sanger_id) %>%
  display_table()

joined %>% 
  mutate(
    has_depmap_id = !is.na(depmap_id),
    has_sanger_id = !is.na(sanger_model_id),
    has_ccle_id_broad = !is.na(ccle_id_broad),
    has_ccle_id_sanger = !is.na(ccle_id_sanger),
    has_cell_line_name_broad = !is.na(cell_line_name_broad),
    has_cell_line_name_sanger = !is.na(cell_line_name_sanger)
  ) %>% 
  count(has_depmap_id, has_sanger_id, has_ccle_id_broad, has_ccle_id_sanger, has_cell_line_name_broad, has_cell_line_name_sanger) %>% 
  arrange(!has_depmap_id, !has_sanger_id, !has_ccle_id_broad, !has_ccle_id_sanger, !has_cell_line_name_broad, !has_cell_line_name_sanger) %>%
  display_table()
```


# Add in GDSC cell line metadata
```{r}
gdsc_cell_lines_metadata = 
  read_tsv("../gdsc_bq_etl/large_files/harmonized/cell_lines_metadata.tsv") %>% 
  select(
    cell_line_name, 
    cosmic_sample_id, 
    sanger_model_id = cell_model_passports_id
  ) %>% 
  mutate(across(everything(), toupper)) %>% 
  mutate(in_gdsc = T) %>% 
  distinct()
```

```{r}
joined2 = 
  full_join(
    joined,
    gdsc_cell_lines_metadata,
    by = "sanger_model_id"
  ) %>% 
  mutate(in_gdsc = ifelse(is.na(in_gdsc), F, T)) %>% 
  select(depmap_id, sanger_model_id, in_gdsc) %>% 
  distinct()


```
```{r}
joined2 %>%
  display_table()
```





