---
title: "Parse Drugbank XML"
author: "Ariel Balter"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Knitr
```{r, include=F}
knitr::opts_chunk$set(
  warning=F,
  message=F,
  echo=F,
  error=F,
  number_section=T,
  include=F
)
```

# Libraries
```{r, include=F}
# Load Libraries
library(vroom)
library(tidyverse)
library(ggplot2)
library(kableExtra)
library(here)
library(jsonlite)
library(reticulate)
use_python("/home/balter/conda/bin/python")
use_condaenv("base")
```


```{bash}
mkdir -p DrugBank/large_file
```


# Download the Schema (XSD)
```{bash}
curl \
  -Lfv \
  -o DrugBank/large_files/drugbank_xsd.zip \
  -u ariel.balter@providence.org:ThreeSix@^3 \
  https://go.drugbank.com/docs/drugbank.xsd
```

# Download the full database
```{bash}
curl \
  -Lfv \
  -o large_files/drugbank_full_database.zip \
  -u ariel.balter@providence.org:ThreeSix@^3 \
  https://go.drugbank.com/releases/5-1-9/downloads/all-full-database
```


```{bash}
which python
```


```{python}
import xmltodict
import json
import os

```

```{bash}
pwd
```

```{r}
list.files()
```



```{python}
os.listdir()

# with open("large_files/full_drugbank_database.xml") as f:
#     xml = xmltodict.parse(f.readlines())

with open("large_files/full_drugbank_database.xml", "rb") as f:
  xml = f.read()

print("parsing to dict")
d = xmltodict.parse(xml, process_namespaces=False)

with open("large_files/full_drugbank_database.json", "w") as f:
  json.dump(d, f, indent=4, sort_keys=True)

drugs = d['drugbank']['drug']

with open("large_files/drugbank_drugs.json", "w") asf:
  json.dump(drugs, f, indent=4, sort_keys=True)
```

```{python}
def depth(d):
  if isinstance(d, dict):
    return 1 + (max(map(depth, d.values())) if d else 0)
  return 0

depth(ddd)
```




https://stackoverflow.com/a/68754233/188963



dict_keys(['@type', '@created', '@updated', 'drugbank-id', 'name', 'description', 'cas-number', 'unii', 'state', 'groups', 'general-references', 'synthesis-reference', 'indication', 'pharmacodynamics', 'mechanism-of-action', 'toxicity', 'metabolism', 'absorption', 'half-life', 'protein-binding', 'route-of-elimination', 'volume-of-distribution', 'clearance', 'classification', 'salts', 'synonyms', 'products', 'international-brands', 'mixtures', 'packagers', 'manufacturers', 'prices', 'categories', 'affected-organisms', 'dosages', 'atc-codes', 'ahfs-codes', 'pdb-entries', 'fda-label', 'msds', 'patents', 'food-interactions', 'drug-interactions', 'sequences', 'experimental-properties', 'external-identifiers', 'external-links', 'pathways', 'reactions', 'snp-effects', 'snp-adverse-drug-reactions', 'targets', 'enzymes', 'carriers', 'transporters'])


