---
title: "CCLE Metadata Files"
author: "Ariel Balter"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: yes
    theme: lumen
    fig_width: 8
    fig_height: 6
    fig_caption: yes
    code_folding: hide
---


```{r}
library(tidyverse)
library(bigrquery)
library(jsonlite)
library(openxlsx)
```


```{r}
knitr::opts_chunk$set(
  attr.source='.numberLines',
  attr.output='style="max-height: 300px;"'
)
```


# Initialize BQ, datasets, and tables
```{r}
bq_auth()

bq_create = function(
    obj_type, 
    project_name="", 
    dataset_name="", 
    table_name=""
  )
{
  # obj_type = "table"
  # project = "katana-providence"
  # dataset_name = dataset_bq
  # table_name = "test"
  
  print(obj_type)
  print(project_name)
  print(dataset_name)
  print(table_name)
  
  print(typeof(dataset_name))
  print(typeof(table_name))
  
  if (obj_type == "dataset")
  {
    print("doing dataset")
    dataset_bq = bq_dataset(project_name, dataset_name)
    print(dataset_bq)

    if (bq_dataset_exists(dataset_bq))
    {
      print("deleting dataset")
      bq_dataset_delete(dataset_bq, delete_contents = T)
    }
    
    bq_dataset_create(dataset_bq)
    
    return(dataset_bq)
    
  } else if(obj_type == "table")
  {
    print("doing table")
    table_bq = bq_table(dataset_name, table_name)
    
    if (bq_table_exists(table_bq))
    {
      bq_table_delete(table_bq)
    }
    
    bq_table_create(table_bq)
    
    return(table_bq)
    
  }
}

project_name = "katana-providence"
dataset_name = "ccle_raw"

dataset_bq = bq_create("dataset", project_name, dataset_name)
```

# Download Files
```{r, label="load wide tables"}
model_list = 
  read_csv("https://cog.sanger.ac.uk/cmp/download/model_list_latest.csv.gz") %>% 
  rename(
    sanger_model_id = model_id,
    sanger_sample_id = sample_id,
    sanger_patient_id = patient_id
  )

sample_info = read_csv("https://ndownloader.figshare.com/files/35020903")

# all_mutations = 
#   read_csv("https://cog.sanger.ac.uk/cmp/download/mutations_all_latest.csv.gz")  %>%
#   rename(
#     sanger_gene_id = gene_id,
#     sanger_model_id = model_id
#     )

# A list of cancer driver mutations present in all sequenced models.
cancer_driver_mutations = 
  read_csv("https://cog.sanger.ac.uk/cmp/download/driver_mutations_latest.csv") %>% 
  rename(sanger_gene_id = gene_id)

# A list of all mutations present in all sequenced models.
cancer_mutations_summary = 
  read_csv("https://cog.sanger.ac.uk/cmp/download/mutations_summary_latest.csv.gz") %>% 
  rename(sanger_gene_id = gene_id)

cancer_driver_genes = 
  read_csv("https://cog.sanger.ac.uk/cmp/download/driver_genes_latest.csv") %>% 
  rename(sanger_gene_id = gene_id)

gene_identifiers = 
  read_csv("https://cog.sanger.ac.uk/cmp/download/gene_identifiers_latest.csv.gz") %>% 
  rename(sanger_gene_id = gene_id)

cancer_predisposition_variants = 
  read_csv("https://cog.sanger.ac.uk/cmp/download/cancer_predisposition_variants_latest.csv") %>% 
  rename(sanger_gene_id = gene_id)

# for model list https://cellmodelpassports.sanger.ac.uk/documentation/models/glossary

metadata_files = list(
  model_list = list(data = model_list),
  sample_info = list(
    data = sample_info, 
    schema = "sample_info_schema.xlsx"
  ),
  # cancer_driver_mutations = list(
  #   data = cancer_driver_mutations, 
  #   schema = "driver_mutations_schema.xlsx"
  # ),
  cancer_mutations_summary = list(
    data = cancer_mutations_summary,
    schema = "ccla_mutations_schema.xlsx"
  ),
  cancer_driver_genes = list(
    data = cancer_driver_genes,
    schema = ""
  ),
  gene_identifiers = list(data = gene_identifiers))
```


# Upload to BQ
```{r, lable="metadata"}

for (table_name in names(metadata_files))
{
  print(table_name)
  
  bq_perform_upload(
    bq_table(dataset_bq, table_name),
    values = metadata_files[[table_name]],
    create_disposition = "CREATE_IF_NEEDED",
    write_disposition = "WRITE_TRUNCATE"
  )
}
```

# DepMap Sample Info
https://depmap.org/portal/download/

```{r}
sample_info = read_csv("https://ndownloader.figshare.com/files/35020903")


```




# Driver Genes 
There is an HTML table here:

https://depmap.sanger.ac.uk/documentation/datasets/gene-annotation-mapping/

The driver mutations table is the union of three datasets:

* <a href="https://www.cancergenomeinterpreter.org/mutations">CGI catalog</a> of validated oncogenic mutations

* <a href="https://www.cancerhotspots.org/swagger-ui.html">MSKCC Hotspots</a> version 2 Includes a set of in-frame indels unique to this dataset

* <a href="https://www.intogen.org/boostdm/downloads">BoostDM dataset</a> A new machine learning based classification of mutation in the intOGen cohorts

The CGI and MSKCC datasets have chromosome coordinates recorded on <a href="https://www.ncbi.nlm.nih.gov/grc/human#previous_major_patch">GRCh37</a>, and were remapped to <a href="https://www.ncbi.nlm.nih.gov/grc/human#current">GRCh38</a> using UCSC’s <a href="https://hgdownload.cse.ucsc.edu/goldenPath/hg19/liftOver/hg19ToHg38.over.chain.gz">hg19 to hg38 chain file</a> and <a href="https://genome-store.ucsc.edu/">liftOver</a> utility. BoostDM is provided on GRCh38.

Driver mutations schema copied from the table on that page.

```{r}
model_dataset_availability = read_csv("https://cog.sanger.ac.uk/cmp/download/model_dataset_availability_latest.csv.gz")

gene_list = read_csv("https://cog.sanger.ac.uk/cmp/download/gene_identifiers_latest.csv.gz")

ccle_mutations = read_csv("https://ndownloader.figshare.com/files/34989940")
```


